---
title: "tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bios625)
```

### Simulate Data  

SNP datasets simulated with this function sim_SNP(). (But they are rather unrealistic, so do not use them forany purpose other than package tutorial.)

```{r}
#Generate SNP dataset
SNP_example=sim_SNP(method="SNP")

#List of population grouping
SNP_example=sim_SNP(method="label")

#List of both
SNP_example=sim_SNP()
```

Overview of the dataset:

```{r}
head(SNP_example)[['SNP_example']][1:6,1:10]
head(SNP_example)[['label']][1:6]
```

Separate the dataset by population groups:
```{r}
SNP=data.prep(SNP_example[['SNP_example']],SNP_example[['label']])
```

Overview of the dataset:

```{r}
SNP$'1'[1:6,1:10]
SNP$'2'[1:6,1:10]
SNP$'2'[1:6,1:10]
```

Compute Fst list:
```{r}
#default: average
FST(SNP$'1',SNP$'3')
```

```{r}
#average
FST(SNP$'2',SNP$'3',method='average')
```

```{r}
#list of all Fst of the all alleles
Fst=FST(SNP$'2',SNP$'3',method='list')
head(Fst)
```


## Compare with pcakage KRIS

The function fst.hudson() in package KRIS does the population group separation within an apply loop used to calculate Fst. FST() of the 625 package pre divided the population into groups, and uses colSum() rather than a loop to calculate each Fst value. so it is more efficient.

```{r}
#setup of fst.hudson package, population group index
idx1 <- which(SNP_example[['label']]== '1')
idx2 <- which(SNP_example[['label']]== '2')

#check result
testthat::expect_equal( FST(SNP$'1',SNP$'2'),
              KRIS::fst.hudson(SNP_example[[1]],idx1,idx2))
testthat::expect_equal( FST(SNP$'1',SNP$'2',method='list'),
                          KRIS::fst.each.snp.hudson(SNP_example[[1]],idx1,idx2))
#compare time
run=bench::mark(FST(SNP$'1',SNP$'2'),
              KRIS::fst.hudson(SNP_example[[1]],idx1,idx2))
ggplot2::autoplot(run)


```


